\thetitle

### 2008-01-17 models.tex
More work on the models section.

\begin{tabular}{|c|l|p{5in}|}
\multicolumn{1}{c}{\textbf{Behavior}} &
\multicolumn{1}{c}{\textbf{Model}} &
\multicolumn{1}{c}{\textbf{Description}} \\
\hline
\multirow{3}[13]{*}{Network}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{T}race} &
Mapping of flow end-points to wireless nodes taken directly from trace data. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{M}arginal} &
Use marginal empirical distributions from trace. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{S}ynthetic} &
Use synthetic distributions \textit{\`a la} Hern\'andez-Campos~\cite{Hernandez06:wlan-traffic}. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{U}niform} &
Both end-points are uniformly randomly chosen from all of the wireless nodes. \\
\hline
\multirow{2}[11]{*}{Flow}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{T}race} &
Each flow has the actual start time, end time, total data sent, and number of packets from the trace. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}{\textbf{P}roportional} &
Byte count taken from trace; sizes and intervals are CBR or VBR, on a per-flow basis. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}{\textbf{U}niform} &
All flows have the same duration, volume, packet rate, and data rate as the trace average. \\
\hline
\multirow{2}[11]{*}{Packet}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{T}race} &
The sequence of packet sizes and inter-transmission intervals is taken directly from the trace. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{S}ample} &
Packets sizes and intervals are randomly drawn from per-flow empirical distributions (VBR). \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{U}niform} &
Packets sizes are uniform and the inter-packet interval is constant (CBR). \\
\hline
\end{tabular}

>>>>

\begin{tabular}{|c|l|p{5in}|}
\multicolumn{1}{c}{\textbf{Behavior}} &
\multicolumn{1}{c}{\textbf{Model}} &
\multicolumn{1}{c}{\textbf{Description}} \\
\hline
\multirow{3}[13]{*}{Network}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{T}race} &
Mapping of flow end-points to wireless nodes taken directly from trace data. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{M}arginal} &
Use marginal empirical distributions from trace. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{S}ynthetic} &
Use synthetic distributions \textit{\`a la} Hern\'andez-Campos~\textit{et~al.}~\cite{Hernandez06:wlan-traffic}. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{U}niform} &
Both end-points are uniformly randomly chosen from all of the wireless nodes. \\
\hline
\multirow{2}[11]{*}{Flow}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{T}race} &
Each flow has the actual start time, end time, total data sent, and number of packets from the trace. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}{\textbf{P}roportional} &
Byte count taken from trace; sizes and intervals are CBR or VBR, on a per-flow basis. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}{\textbf{U}niform} &
All flows have the same duration, volume, packet rate, and data rate as the trace average. \\
\hline
\multirow{2}[11]{*}{Packet}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{T}race} &
The sequence of packet sizes and inter-transmission intervals is taken directly from the trace. \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{S}ample} &
Packets sizes and intervals are randomly drawn from per-flow empirical distributions (VBR). \\
\cline{2-3}
\bigstrut & \multirow{1}{*}[-0.05em]{\textbf{U}niform} &
Packets sizes are uniform and the inter-packet interval is constant (CBR). \\
\hline
\end{tabular}

### 2008-01-17 paper.tex
More work on the models section.

\begin{table*}
\begin{center}\small
\input{models.tex}
\caption{The three orthogonal levels of traffic behavior, and the traffic models used for each level.}\label{tab:traffic-models}
\end{center}
\vspace{-1.75em}
\end{table*}

@@

We use the Qualnet wireless network simulator to perform our experiments. We simulate a stationary multi-hop 802.11b network using the Ad hoc On-demand Distance Vector (\caps{AODV}) routing protocol~\cite{rfc:aodv}, with nodes placed randomly in a square field with sides of 150 meters.
In addition to the active nodes corresponding to trace \caps{IP}s, half as many passive ``infrastructure'' nodes are added to each simulation: these nodes initiate no data and simply serve as additional network relays. Our simulations resemble multi-hop mesh networks of the kind that are increasingly studied and deployed for delivery of broadband access in residential, corporate and conference settings. We do not attempt to reproduce the physical environment of the original wireless network, nor do we simulate mobility. The only aspect of the original network's behavior that is reproduced is the total pattern of network-wide traffic.

There are a number of potential objections to this approach. We use single-hop trace data to drive multi-hop simulations; the physical environment, node mobility, handover behavior, and closed-loop dynamics (including \caps{TCP} feedback) of the original wireless setting are not faithfully reproduced. One must keep in mind, however, that the goal of this research is \textit{not} to understand the conditions of the original network. Rather, we are using the traffic behaviors observed as examples to help us better understand how different types of workload can affect performance metrics. In particular, we aim to understand how real workload compares with common synthetic traffic models. Of course, the reason for such objections is that networking researchers understand that the many aspects of behavior interact with each other in a complex and nearly inextricable manner. However, before we can hope to understand the interaction between workload and other features affecting network behavior, we must study traffic patterns alone, and learn to model them with reasonable accuracy in the absence of additional complicating factors. Accordingly, in this study, we detach application level traffic patterns from the other factors influencing network conditions, and study them in isolation.

@@

We separate our traffic generation models into three orthogonal and nearly independent levels of behavior:
\begin{enumerate}
\item \textbf{Network Behavior}: which nodes communicate with each other, and how frequently; \textit{i.e.} how flow end-points are mapped onto nodes in the network.
\item \textbf{Flow Behavior}: high-level parameters for each flow, including start time, end time, packets sent, bytes sent.
\item \textbf{Packet Behavior}: sizes of individual packets, and the intervals between their transmission.
\end{enumerate}
For each level of traffic behavior, we compare several different behavior models. The different levels of models and variants at each level are listed and described in Table~\ref{tab:traffic-models} and illustrated with specific examples in Figure~\ref{fig:traffic-models}.

>>>>

\begin{table*}
\begin{center}\small
\input{models.tex}
\caption{The three levels of traffic behavior and the traffic models defined for each level.}\label{tab:traffic-models}
\end{center}
\vspace{-1.75em}
\end{table*}

@@

We use the Qualnet wireless network simulator to perform our experiments. We simulate a stationary multi-hop 802.11g network using the Ad hoc On-demand Distance Vector (\caps{AODV}) routing protocol~\cite{rfc:aodv}, with nodes placed randomly in a square field with sides of 150 meters.
In addition to the active nodes corresponding to trace \caps{IP}s, half as many passive ``infrastructure'' nodes are added to each simulation: these nodes initiate no data and simply serve as additional network relays. Our simulations resemble multi-hop mesh networks of the kind that are increasingly studied and deployed for delivery of broadband access in residential, corporate and conference settings. We do not attempt to reproduce the physical environment of the original wireless network, nor do we simulate mobility. The only aspect of the original network's behavior that is reproduced is the total pattern of network-wide traffic.

There are a number of potential objections to this approach. We use single-hop trace data to drive multi-hop simulations; the physical environment, node mobility, handover behavior, and closed-loop dynamics of the original wireless setting are not faithfully reproduced. One must keep in mind, however, that the goal of this research is \textit{not} to understand the conditions of the original network. Rather, we are using the traffic behaviors observed as examples to help us better understand how different types of workload can affect performance metrics. In particular, we aim to understand how real workload compares with common synthetic traffic models. Of course, the reason for such objections is that networking researchers understand that the many aspects of behavior interact with each other in a complex and nearly inextricable manner. However, before we can hope to understand the interaction between workload and other features affecting network behavior, we must study traffic patterns alone, and learn to model them with reasonable accuracy in the absence of additional complicating factors. Accordingly, in this study, we detach application level traffic patterns from the other factors influencing network conditions, and study them in isolation.

@@

We separate traffic generation into three levels of behavior: network, flow and packet. The {network} level determines properties pertaining to nodes and their interaction. For example, it determines which nodes communicate with each other, and how frequently. This level can be viewed as assigning network nodes to the source and destination roles of flows, as well as determining per-node properties which affect lower level models. The next level of behavior determines overall properties of flows. This includes each flow's start time, duration, packet count and total data volume. Finally, the packet level of behavior determines the individual sizes of packets and the durations of intervals between their transmission. Despite the concreteness of the examples mentioned here, there are also less obvious, abstract properties that models at the network and flow levels determine. Flow models, for example, dictate the distribution of each flow's packet sizes, a property which can be used to determine individual packet sizes, but which is not itself specific to any particular packet.

%\subsubsection*{Network Behavior Models}

For each level of behavior we define a spectrum of models. The gold standard of realism at each level is provided by the ``trace'' model for that level, which replicates the node, flow or packet behavior of the original trace exactly. The most rudimentary model at each level is the ``uniform'' model, which treats all entities identically---be they nodes, flows or packets. Despite being so simplistic (or perhaps because of it), the uniform behavior models are by far the most commonly for workload generation in experiments. Between these extremes on the spectrum of realism, we consider one or two models with intermediate degrees of complexity and fidelity. At the packet level, we define the ``sample'' model, which is in essence a per-flow variable bit-rate (\caps{VBR}) model: each packet's size is determined by randomly sampling a value from the packet size distribution given by the flow model; the interval until the next packet is then sampled from the given interval distribution.\footnote{Note that the uniform packet model (\textit{i.e.} {\scriptsize{CBR}}), can be considered to be a special case of the sample model where the size and time distributions are both constant.} The intermediate model at the flow level is designated as ``proportional'': the distributions of inter-packet ...

%\begin{enumerate}
%\item \textbf{Network Behavior}: which nodes communicate with each other, and how frequently; \textit{i.e.} how flow end-points are mapped onto nodes in the network.
%\item \textbf{Flow Behavior}: high-level parameters for each flow, including start time, end time, packets sent, bytes sent.
%\item \textbf{Packet Behavior}: sizes of individual packets, and the intervals between their transmission.
%\end{enumerate}

%The different levels of models and variants at each level are listed and described in Table~\ref{tab:traffic-models} and illustrated with specific examples in Figure~\ref{fig:traffic-models}.
